---
title: "Maize Height Models"
subtitle: "Different Maize Models"
author: "Metodi Latinov"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-tools: true
    theme: cosmo
    fig-width: 8
    fig-height: 5
    fig-cap-location: bottom
execute:
  echo: true
  warning: false
  message: false
  cache: false
knitr:
  opts_chunk:
    fig.align: center
---

```{r setup}
cat("Working directory:", getwd(), "\n")
cat("_targets dir exists:", dir.exists("_targets"), "\n")
cat("_targets.R exists:", file.exists("_targets.R"), "\n")
cat("Files here:", paste(head(list.files()), collapse = ", "), "\n")

# Libraries
library(targets)
library(patchwork)
library(tidyverse)
library(stanviz)
library(ggdist)
library(tidyr)
library(posterior)
library(kableExtra)

# Load the cleaned data from the pipeline and models
tar_load(model_data_maize_heigh)

## M5 Hierarchical cell-means, pots pooled within treatment 
tar_load(m5_maize_model)
tar_load(m5_maize_model_pp)

## M6 Hierarchical Golmez
tar_load(m6_maize_model)
tar_load(m6_maize_model_pp)

```

# Hierarchical cell-means pots pooled within treatment 

## Posterior Predictive Checks

```{r}
m5_ppc_dens <- plot_ppc_by_group(
    model         = m5_maize_model,
    yrep_var      = "h_rep",
    group         = model_data_maize_heigh$day_idx,
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
) + coord_cartesian(xlim = c(0, 80))

m5_ppc_mean <- plot_ppc_stat(
    model    = m5_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "mean"
) + coord_cartesian(xlim = c(15, 35))

m5_ppc_sd <- plot_ppc_stat(
    model    = m5_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m5_ppc_mean + m5_ppc_sd) / m5_ppc_dens +
    plot_annotation(
        title = "Posterior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

## Predictive error scatter
```{r}
plot_ppc_error_scatter(
    model = m5_maize_model,
    y = model_data_maize_heigh$plant_height,
    yrep_var = "h_rep",
    x_label = "Plant Height"
) + labs(
        title = "Posterior Predictive Errors",
        y     = "Predicted Error = Y - E[Y_rep]") +
    theme(
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
    )
```

## Inference

### Cell-level summaries Intervals 

```{r}
day_7_f <- plot_coef_forest(
    model = m5_maize_model,
    pars = c("median_height[1,1]","median_height[2,1]","median_height[3,1]"),
    labels = c(
        "median_height[1,1]" = "Control",
        "median_height[2,1]" = "SHAM",
        "median_height[3,1]" = "EMS"
    ),
    null_line = 10.4
) + labs(
    title = "Posterior Maize Height Day 7",
     x = "Posterior Height (cm)"
    )

day_14_f <- plot_coef_forest(
    model = m5_maize_model,
    pars = c("median_height[1,2]","median_height[2,2]","median_height[3,2]"),
    labels = c(
        "median_height[1,2]" = "Control",
        "median_height[2,2]" = "SHAM",
        "median_height[3,2]" = "EMS"
    ),
    null_line = 24.3
) + labs(
    title = "Posterior Maize Height Day 14",
    x = "Posterior Height (cm)"
    )

day_20_f <- plot_coef_forest(
    model = m5_maize_model,
    pars = c("median_height[1,3]","median_height[2,3]","median_height[3,3]"),
    labels = c(
        "median_height[1,3]" = "Control",
        "median_height[2,3]" = "SHAM",
        "median_height[3,3]" = "EMS"
    ),
    null_line = 30.6
) + labs(
    title = "Posterior Maize Height Day 20",
    x = "Posterior Height (cm)"
    )

day_27_f <- plot_coef_forest(
    model = m5_maize_model,
    pars = c("median_height[1,4]","median_height[2,4]","median_height[3,4]"),
    labels = c(
        "median_height[1,4]" = "Control",
        "median_height[2,4]" = "SHAM",
        "median_height[3,4]" = "EMS"
    ),
    null_line = 35.8
)+ labs(
    title = "Posterior Maize Height Day 27",
    x = "Posterior Height (cm)"
    )

## Combine the plots 
(day_7_f + day_14_f) / (day_20_f + day_27_f) +
    plot_annotation(
        title = "Posterior Maize Height Day by Day Comparison",
        theme = theme(
            plot.title = element_text(
                size = 20,
                face = "bold",
                hjust = 0.5,
            )
        )
)
```

### Pairwise contrasts per day

```{r}
plot_param_halfeye(
    model = m5_maize_model,
    pars  = c(
        "log_diff_emf_vs_control[1]",
        "log_diff_emf_vs_control[2]",
        "log_diff_emf_vs_control[3]",
        "log_diff_emf_vs_control[4]"
    ),
    labels = c(
        "log_diff_emf_vs_control[1]" = "Day 7",
        "log_diff_emf_vs_control[2]" = "Day 14",
        "log_diff_emf_vs_control[3]" = "Day 20",
        "log_diff_emf_vs_control[4]" = "Day 27"
    )
) + labs(
    title = "Posterior Log Difference EMS - Control by Days",
    x     = "Posterior Log Difference EMS - Control" 
    ) + 
    theme(plot.title = element_text(size = 20, face = "bold"))

```

### Probability-of-direction indicators 

```{r}
pd_emf_control <- m5_maize_model$draws("prob_emf_vs_control_positive", format = "draws_matrix") %>%
    colMeans()
pd_emf_sham <- m5_maize_model$draws("prob_emf_vs_sham_positive", format = "draws_matrix") %>%
    colMeans()
pd_sham_control <- m5_maize_model$draws("prob_sham_vs_control_positive", format = "draws_matrix") %>%
    colMeans()

# Stack the three contrasts in a single faceted plot
pd_long <- tibble::tibble(
    day = rep(c(7, 14, 20, 27), 3),
    contrast = rep(c("EMF vs Control", "EMF vs Sham", "Sham vs Control"), each = 4),
    pd = c(pd_emf_control, pd_emf_sham, pd_sham_control)
    )

pd_long <- pd_long %>%
    mutate(direction = ifelse(pd > 0.5, "positive", "negative"))

ggplot(pd_long, aes(x = factor(day), y = pd)) +
    geom_segment(aes(
        xend = factor(day), 
        y = 0.5, 
        yend = pd, 
        color = pd),
        linewidth = 6
    ) +
    geom_point(aes(color = pd), size = 3) +
    geom_hline(yintercept = 0.5, color = "gray50", linewidth = 0.4) +
    scale_color_gradient2(
        low = "#D85A30",      
        mid = "gray80",       
        high = "#1D9E75",     
        midpoint = 0.5,
        limits = c(0, 1)
    ) +
    facet_wrap(~ contrast) +
    coord_cartesian(ylim = c(0, 1)) +
    labs(
        x = "Day", 
        y = "Posterior probability contrast > 0",
        title = "Probability-of-direction across all three contrasts"
    ) +
    theme_bayes(base_size = 14) +
    theme(legend.position = "none")
```

### Effect posterior vs. Region Of Practical Equivalence 
```{r}
# Pull the actual contrast draws (not just the ROPE indicator)
contrast_draws <- m5_maize_model$draws("log_diff_emf_vs_control", format = "draws_matrix")

# For each day (column), compute the three regions
rope_summary <- tibble::tibble(
    day = c(7, 14, 20, 27),
    below = colMeans(contrast_draws < -0.05),
    in_rope = colMeans(abs(contrast_draws) < 0.05),
    above = colMeans(contrast_draws > 0.05)
) %>%
    tidyr::pivot_longer(c(below, in_rope, above), names_to = "region", values_to = "prob") |>
    dplyr::mutate(region = factor(region, levels = c("below", "in_rope", "above")))

ggplot(rope_summary, aes(x = factor(day), y = prob, fill = region)) +
    geom_col(width = 0.7) +
    scale_fill_manual(
        values = c(
            below   = "#D85A30",   # coral: effect clearly negative
            in_rope = "gray70",      # gray: effectively null
            above   = "#1D9E75"    # teal: effect clearly positive
        ),
        labels = c("< -ROPE", "in ROPE", "> +ROPE")
    ) +
    coord_cartesian(ylim = c(0, 1)) +
    labs(
        x = "Day",
        y = "Posterior probability",
        fill = "Region",
        title = "EMF vs Control: posterior mass by region (ROPE = ±0.05 log-units)"
    ) +
    theme_bayes(base_size = 14)

```

###  Variance decomposition

#### R2
```{r}
full_r2 <- plot_param_halfeye(
    model = m5_maize_model,
    pars  = c("bayes_R2_full"),
    labels = c("bayes_R2_full" = "Bayesian R2 Full Model"),
    null_line = 0.684
) + labs(
    title = "Bayesian Full Model R2",
    x = "R2"
    )

population_r2 <- plot_param_halfeye(
    model = m5_maize_model,
    pars  = c("bayes_R2_population"),
    labels = c("bayes_R2_population" = "Bayesian Population R2 Model"),
    null_line = 0.669
) + labs(
    title = "Bayesian Population Model R2",
    x = "R2"
    )

contribution_r2 <- plot_param_halfeye(
    model = m5_maize_model,
    pars  = c("bayes_R2_pot_contribution"),
    labels = c("bayes_R2_pot_contribution" = "Bayesian Pot Contribution R2 Model"),
    null_line = 0.0149
) + labs(
    title = "Bayesian Pot Contribution Model R2",
    x = "R2")

(full_r2 + population_r2) / contribution_r2 +
    plot_annotation(
        title = "Bayesian R2 Variance Decomposition",
        theme = theme(
            plot.title = element_text(
                size = 20,
                face = "bold",
                hjust = 0.5
            ) 
        ) 
    ) 

```

### Variance decomposition across hierarchical levels
```{r}
plot_variance_decomposition(
    model = m5_maize_model,
    sd_vars = c(
        "Between pots (Control)" = "zeta_t[1]",
        "Between pots (Sham)"    = "zeta_t[2]", 
        "Between pots (EMF)"     = "zeta_t[3]",
        "Within pots, day 7"     = "sigma_d[1]",
        "Within pots, day 14"    = "sigma_d[2]",
        "Within pots, day 20"    = "sigma_d[3]",
        "Within pots, day 27"    = "sigma_d[4]"
    )
) + labs(
    title = "Variance Decomposition Within pots and Between Pots",
    x = "Proportion of Total Variance"
    )+ theme(
        plot.title = element_text(
            size = 20, 
            face = "bold",
            hjust = 0.5
        ),
        plot.subtitle = element_blank()
    )
```

# Hierarchical Gompertz, per-pot curves pooled within treatment 

## Posterior Predictive Checks

```{r}
m6_ppc_dens <- plot_ppc_dens(
    model         = m6_maize_model,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
) + coord_cartesian(xlim = c(0, 80))

m6_ppc_mean <- plot_ppc_stat(
    model    = m6_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "mean"
) + coord_cartesian(xlim = c(15, 35))

m6_ppc_sd <- plot_ppc_stat(
    model    = m6_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m6_ppc_mean + m5_ppc_sd) / m6_ppc_dens +
    plot_annotation(
        title = "Posterior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

## Predictive error scatter
```{r}
plot_ppc_error_scatter(
    model = m6_maize_model,
    y = model_data_maize_heigh$plant_height,
    yrep_var = "h_rep",
    x_label = "Plant Height"
) + labs(
        title = "Posterior Predictive Errors",
        y     = "Predicted Error = Y - E[Y_rep]") +
    theme(
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
    )
```

## Infrence 

### Population-level natural-scale parameters
```{r}
# Pull the population-level parameters as a draws_df
pop_draws <- m6_maize_model$draws(c("A_pop", "K_pop", "c_t"), format = "draws_df")

# Reshape with reordered treatment factor levels for visual clarity
pop_long <- pop_draws %>%
    select(-.chain, -.iteration, -.draw) %>%
    pivot_longer(
        everything(),
        names_to = c("parameter", "treatment_idx"),
        names_pattern = "(A_pop|K_pop|c_t)\\[(\\d+)\\]",
        values_to = "value"
    ) %>%
    mutate(
        treatment = factor(
            treatment_idx,
            levels = c("2", "3", "1"),
            labels = c("Sham", "EMF", "Control")
        ),
        parameter = factor(parameter, levels = c("A_pop", "K_pop", "c_t"),
                          labels = c("Asymptote A (cm)",
                                     "Growth rate K (per day)",
                                     "Inflection point τ (days)"))
    )

# Compute Control's posterior median per parameter for the reference line
control_medians <- pop_long %>%
    filter(treatment == "Control") %>%
    group_by(parameter) %>%
    summarize(control_median = median(value), .groups = "drop")

# Consistent palette
treatment_colors <- c(
    "Control" = "#2E5984",
    "Sham"    = "#B8860B",
    "EMF"     = "#7B2D8E"
)

# Build the plot — point + interval style, with Control reference line
plot_population_params <- ggplot(pop_long, aes(y = treatment, x = value, color = treatment)) +
    geom_vline(
        data = control_medians,
        aes(xintercept = control_median),
        linetype = "dashed",
        color = "gray40",
        linewidth = 0.5
    ) +
    stat_pointinterval(
        .width = c(0.5, 0.9),
        point_size = 3,
        interval_size_range = c(0.6, 1.8)
    ) +
    facet_wrap(~ parameter, scales = "free_x", nrow = 1) +
    scale_color_manual(values = treatment_colors, guide = "none") +
    labs(
        x = NULL,
        y = NULL,
        title = "Population-level Gompertz curve parameters by treatment",
        subtitle = "Point = posterior median; thick line = 50% interval; thin line = 90% interval. Dashed line shows Control median for reference."
    ) +
    theme_paper +
    theme(
        strip.text = element_text(face = "bold", size = 11),
        panel.spacing = unit(1, "lines")
    )
plot_population_params
``` 

#### Table
```{r}
# Pull summary statistics for the population parameters
pop_summary <- m6_maize_model$summary(
    variables = c("A_pop", "K_pop", "c_t"),
    posterior::default_summary_measures()
)

# Build the table: parameter rows, treatment columns, formatted "median [q5, q95]"
table_data <- pop_summary %>%
    select(variable, median, q5, q95) %>%
    mutate(
        parameter = case_when(
            grepl("A_pop", variable) ~ "Асимптота A (cm)",
            grepl("K_pop", variable) ~ "Скорост на растеж K (на ден)",
            grepl("c_t",   variable) ~ "Момент на инфлексия τ (дни)"
        ),
        treatment_idx = as.integer(gsub("[^0-9]", "", variable)),
        treatment = factor(
            treatment_idx,
            levels = c(1, 3, 2),      # Control, EMF, Sham — same order as plot
            labels = c("Контрол", "EMF", "Sham")
        ),
        # Decimal precision matched to parameter scale
        formatted = case_when(
            grepl("A_pop", variable) ~ sprintf("%.1f [%.1f, %.1f]", median, q5, q95),
            grepl("K_pop", variable) ~ sprintf("%.3f [%.3f, %.3f]", median, q5, q95),
            grepl("c_t",   variable) ~ sprintf("%.2f [%.2f, %.2f]", median, q5, q95)
        )
    ) %>%
    select(parameter, treatment, formatted) %>%
    pivot_wider(names_from = treatment, values_from = formatted) %>%
    # Preserve the row order: A, K, tau
    mutate(parameter = factor(
        parameter,
        levels = c("Асимптота A (cm)",
                   "Скорост на растеж K (на ден)",
                   "Момент на инфлексия τ (дни)")
    )) %>%
    arrange(parameter)

# Render as LaTeX
table_latex <- kbl(
    table_data,
    format = "latex",
    booktabs = TRUE,
    align = c("l", "c", "c", "c"),
    col.names = c("Параметър", "Контрол", "EMF", "Sham"),
    caption = "Параметри на популационната Гомперцова крива по третиране. Стойностите са апостериорни медиани с 90\\% достоверни интервали в скоби.",
    escape = FALSE
) %>%
    kable_styling(
        latex_options = c("hold_position"),
        position = "center"
    ) %>%
    add_header_above(c(" " = 1, "Третиране" = 3))

# Print to console (copy-paste into LaTeX)
cat(table_latex)

```

### Pairwise contrasts
```{r}
a <- plot_param_halfeye(
    model = m6_maize_model,
    pars = c("log_A_emf_vs_control"),
    labels = c("log_A_emf_vs_control" = "Golmez Asymptote")
) + labs(
        title = "Log Difference EMF - Control" 
    )+
    theme(
        plot.title = element_text(size = 10, face = "bold",hjust = 0.5)
)

k <- plot_param_halfeye(
    model = m6_maize_model,
    pars = c("log_K_emf_vs_control"),
    labels = c("log_K_emf_vs_control" = "Golmez Growth Rate")
) + labs(
        title = "Log Difference EMF - Control" 
    )+
    theme(
        plot.title = element_text(size = 10, face = "bold",hjust = 0.5)
)

tau <- plot_param_halfeye(
    model = m6_maize_model,
    pars = c("tau_emf_vs_control"),
    labels = c("tau_emf_vs_control" = "Golmez Inflection point")
) + labs(
        title = "Difference EMF - Control" 
    )+
    theme(
        plot.title = element_text(size = 10, face = "bold",hjust = 0.5)
)

(a + k + tau) + 
    plot_annotation(
        title = "Posterior Contrast of the Golmez Parameters",
        theme = theme(
            plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
        )
    )
```

#### Table 
```{r}
# Pull summary statistics for all nine pairwise contrasts
contrasts_summary <- m6_maize_model$summary(
    variables = c(
        # Asymptote contrasts
        "log_A_emf_vs_control", "log_A_emf_vs_sham", "log_A_sham_vs_control",
        # Rate contrasts
        "log_K_emf_vs_control", "log_K_emf_vs_sham", "log_K_sham_vs_control",
        # Inflection contrasts
        "tau_emf_vs_control",   "tau_emf_vs_sham",   "tau_sham_vs_control"
    ),
    posterior::default_summary_measures()
)

# Build the table: parameter rows, comparison columns
table_contrasts <- contrasts_summary %>%
    select(variable, median, q5, q95) %>%
    mutate(
        parameter = case_when(
            grepl("^log_A", variable) ~ "log A (асимптота)",
            grepl("^log_K", variable) ~ "log K (скорост)",
            grepl("^tau",   variable) ~ "τ (инфлексия, дни)"
        ),
        comparison = case_when(
            grepl("emf_vs_control", variable)  ~ "EMF − Контрол",
            grepl("emf_vs_sham", variable)     ~ "EMF − Sham",
            grepl("sham_vs_control", variable) ~ "Sham − Контрол"
        ),
        # Decimal precision matched to parameter scale
        formatted = case_when(
            grepl("^log_A", variable) ~ sprintf("%.3f [%.3f, %.3f]", median, q5, q95),
            grepl("^log_K", variable) ~ sprintf("%.3f [%.3f, %.3f]", median, q5, q95),
            grepl("^tau",   variable) ~ sprintf("%.2f [%.2f, %.2f]", median, q5, q95)
        )
    ) %>%
    select(parameter, comparison, formatted) %>%
    pivot_wider(names_from = comparison, values_from = formatted) %>%
    # Order rows: A, K, tau
    mutate(parameter = factor(
        parameter,
        levels = c("log A (асимптота)",
                   "log K (скорост)",
                   "τ (инфлексия, дни)")
    )) %>%
    arrange(parameter) %>%
    # Order columns: EMF-Control (headline), EMF-Sham, Sham-Control
    select(parameter, `EMF − Контрол`, `EMF − Sham`, `Sham − Контрол`)

# Render as LaTeX
table_latex_contrasts <- kbl(
    table_contrasts,
    format = "latex",
    booktabs = TRUE,
    align = c("l", "c", "c", "c"),
    col.names = c("Параметър", "EMF − Контрол", "EMF − Sham", "Sham − Контрол"),
    caption = "Двойни контрасти за параметрите на Гомперцовата крива. Контрастите за $\\log A$ и $\\log K$ са в логаритмичен мащаб; контрастът за $\\tau$ е в дни. Стойностите са апостериорни медиани с 90\\% достоверни интервали в скоби.",
    escape = FALSE
) %>%
    kable_styling(
        latex_options = c("hold_position"),
        position = "center"
    ) %>%
    add_header_above(c(" " = 1, "Сравнение между третирания" = 3))

# Print to console
cat(table_latex_contrasts)
```

### Probability-of-direction indicators 
```{r}
# Pull the three probability-of-direction indicators
pd_A <- m6_maize_model$draws("prob_log_A_emf_vs_control_positive", format = "draws_matrix") |>
    mean()
pd_K <- m6_maize_model$draws("prob_log_K_emf_vs_control_positive", format = "draws_matrix") |>
    mean()
pd_T <- m6_maize_model$draws("prob_tau_emf_vs_control_positive", format = "draws_matrix") |>
    mean()

# Build the tibble: one row per contrast
pd_df <- tibble::tibble(
    parameter = c("Asymptote (log A)", "Rate (log K)", "Inflection (τ)"),
    pd = c(pd_A, pd_K, pd_T)
) |>
    dplyr::mutate(
        parameter = factor(parameter, levels = parameter),  # preserve order
        direction = ifelse(pd > 0.5, "positive", "negative")
    )

ggplot(pd_df, aes(x = parameter, y = pd)) +
    geom_segment(
        aes(xend = parameter, y = 0.5, yend = pd, color = direction),
        linewidth = 8
    ) +
    geom_point(aes(color = direction), size = 4) +
    geom_hline(yintercept = 0.5, color = "gray50", linewidth = 0.4) +
    geom_hline(yintercept = c(0.05, 0.95), 
               linetype = "dashed", color = "gray70", linewidth = 0.3) +
    scale_color_manual(values = c(
        "positive" = "#1D9E75",
        "negative" = "#D85A30"
    )) +
    coord_cartesian(ylim = c(0, 1)) +
    labs(
        x = "Gompertz parameter",
        y = "P(EMF > Control)",
        title = "Probability-of-direction: EMF vs Control on each curve parameter"
    ) +
    theme_bayes(base_size = 14) +
    theme(legend.position = "none")
```

### Effect posterior vs. Region Of Practical Equivalence 
```{r}
# Pull the three scalar contrast draws
draws_A <- m6_maize_model$draws("log_A_emf_vs_control", format = "draws_matrix")
draws_K <- m6_maize_model$draws("log_K_emf_vs_control", format = "draws_matrix")
draws_T <- m6_maize_model$draws("tau_emf_vs_control",   format = "draws_matrix")

# ROPE widths — different per parameter, matching the scale of each
rope_A <- 0.05
rope_K <- 0.10
rope_T <- 0.5

# Build the region tibble — one row per (parameter × region)
rope_summary <- tibble::tribble(
    ~parameter,         ~rope_width, ~below,                   ~in_rope,                          ~above,
    "Asymptote (log A)", rope_A,     mean(draws_A < -rope_A),  mean(abs(draws_A) < rope_A),       mean(draws_A > rope_A),
    "Rate (log K)",      rope_K,     mean(draws_K < -rope_K),  mean(abs(draws_K) < rope_K),       mean(draws_K > rope_K),
    "Inflection (τ)",    rope_T,     mean(draws_T < -rope_T),  mean(abs(draws_T) < rope_T),       mean(draws_T > rope_T)
) %>%
    dplyr::mutate(parameter = factor(parameter, levels = parameter)) %>%  # preserve order
    tidyr::pivot_longer(c(below, in_rope, above),
                        names_to = "region", values_to = "prob") %>%
    dplyr::mutate(region = factor(region, levels = c("below", "in_rope", "above")))

ggplot(rope_summary, aes(x = parameter, y = prob, fill = region)) +
    geom_col(width = 0.7) +
    scale_fill_manual(
        values = c(
            below   = "#D85A30",   # coral: effect clearly negative
            in_rope = "gray70",    # gray: effectively null
            above   = "#1D9E75"    # teal: effect clearly positive
        ),
        labels = c("< -ROPE", "in ROPE", "> +ROPE")
    ) +
    coord_cartesian(ylim = c(0, 1)) +
    labs(
        x = "Gompertz parameter",
        y = "Posterior probability",
        fill = "Region",
        title = "EMF vs Control: posterior mass by region",
        subtitle = "ROPE widths: log A ±0.05 (~5% asymptote); log K ±0.10 (~10% rate); τ ±0.5 days"
    ) +
    theme_bayes(base_size = 14)
```
###  Variance decomposition

#### R2
```{r}
full_r2 <- plot_param_halfeye(
    model = m6_maize_model,
    pars  = c("bayes_R2_full"),
    labels = c("bayes_R2_full" = "Bayesian R2 Full Model"),
    null_line = 0.684
) + labs(
    title = "Bayesian Full Model R2",
    x = "R2"
    )

population_r2 <- plot_param_halfeye(
    model = m6_maize_model,
    pars  = c("bayes_R2_population"),
    labels = c("bayes_R2_population" = "Bayesian Population R2 Model"),
    null_line = 0.669
) + labs(
    title = "Bayesian Population Model R2",
    x = "R2"
    )

contribution_r2 <- plot_param_halfeye(
    model = m6_maize_model,
    pars  = c("bayes_R2_pot_contribution"),
    labels = c("bayes_R2_pot_contribution" = "Bayesian Pot Contribution R2 Model"),
    null_line = 0.0149
) + labs(
    title = "Bayesian Pot Contribution Model R2",
    x = "R2"
    )

(full_r2 + population_r2) / contribution_r2 +
    plot_annotation(
        title = "Bayesian R2 Variance Decomposition",
        theme = theme(
            plot.title = element_text(
                size = 20,
                face = "bold",
                hjust = 0.5
            ) 
        ) 
    ) 

```

#### Table 
```{r}
# Pull summary stats for the three R² components
r2_summary <- m6_maize_model$summary(
    variables = c("bayes_R2_full", "bayes_R2_population", "bayes_R2_pot_contribution"),
    posterior::default_summary_measures()
)

# Build the table: each row is one R² component
table_r2 <- r2_summary %>%
    select(variable, median, q5, q95) %>%
    mutate(
        label = case_when(
            variable == "bayes_R2_full"             ~ "$R^2_{\\text{пълен}}$ (с информация на ниво саксия)",
            variable == "bayes_R2_population"       ~ "$R^2_{\\text{популация}}$ (само на ниво третиране и ден)",
            variable == "bayes_R2_pot_contribution" ~ "$R^2_{\\text{принос на саксията}}$ (разлика)"
        ),
        formatted = sprintf("%.3f [%.3f, %.3f]", median, q5, q95)
    ) %>%
    mutate(label = factor(label, levels = c(
        "$R^2_{\\text{пълен}}$ (с информация на ниво саксия)",
        "$R^2_{\\text{популация}}$ (само на ниво третиране и ден)",
        "$R^2_{\\text{принос на саксията}}$ (разлика)"
    ))) %>%
    arrange(label) %>%
    select(label, formatted)

# Render as LaTeX
table_latex_r2 <- kbl(
    table_r2,
    format = "latex",
    booktabs = TRUE,
    align = c("l", "c"),
    col.names = c("Компонент", "Медиана [90\\% ДИ]"),
    caption = "Разлагане на дисперсията чрез Байесов $R^2$ за Гомперцовия модел. Стойностите са апостериорни медиани с 90\\% достоверни интервали в скоби.",
    escape = FALSE
) %>%
    kable_styling(
        latex_options = c("hold_position"),
        position = "center"
    )

cat(table_latex_r2)
```
### Curve Prediction Grid 
```{r}
# 1. Extract the grid as a draws array — gets the 3D structure
mu_grid_draws <- m6_maize_model$draws("mu_pop_grid", format = "draws_array")
t_grid_values <- m6_maize_model$draws("t_grid", format = "draws_matrix") |>
    as.numeric() %>%
    unique()  # All draws have the same t_grid since it's deterministic

# 2. Convert to long tibble: one row per (draw, treatment, time)
mu_grid_long <- mu_grid_draws %>%
    posterior::as_draws_df() %>%
    pivot_longer(
        starts_with("mu_pop_grid"),
        names_to = c("treatment", "g"),
        names_pattern = "mu_pop_grid\\[(\\d+),(\\d+)\\]",
        values_to = "height"
    ) %>%
    mutate(
        treatment = factor(treatment, 
                          levels = c("1", "2", "3"),
                          labels = c("Control", "Sham", "EMF")),
        g = as.integer(g),
        t = t_grid_values[g]
    )

# 3. Summarize per (treatment, t): median + intervals
curve_summary <- mu_grid_long %>%
    group_by(treatment, t) %>%
    summarize(
        median = median(height),
        q05 = quantile(height, 0.05),
        q25 = quantile(height, 0.25),
        q75 = quantile(height, 0.75),
        q95 = quantile(height, 0.95),
        .groups = "drop"
    )

# 4. Prep the raw data for overlay
raw_data <- model_data_maize_heigh %>%
    mutate(treatment = factor(treatment, 
                              levels = c(1, 2, 3),
                              labels = c("Control", "Sham", "EMF")))

# 5. Define the palette — distinguishable, color-blind-safe ideally
treatment_colors <- c(
    "Control" = "#2E5984",   # blue
    "Sham"    = "#B8860B",   # ochre
    "EMF"     = "#7B2D8E"    # purple
)

# 6. Build the plot
ggplot() +
    # Ribbon: 90% credible interval (outer)
    geom_ribbon(
        data = curve_summary,
        aes(x = t, ymin = q05, ymax = q95, fill = treatment),
        alpha = 0.15
    ) +
    # Ribbon: 50% credible interval (inner)
    geom_ribbon(
        data = curve_summary,
        aes(x = t, ymin = q25, ymax = q75, fill = treatment),
        alpha = 0.25
    ) +
    # Median curve
    geom_line(
        data = curve_summary,
        aes(x = t, y = median, color = treatment),
        linewidth = 1.2
    ) +
    # Raw data points, jittered slightly on x to reveal overlapping points
    geom_jitter(
        data = raw_data,
        aes(x = day, y = plant_height, color = treatment),
        width = 0.3, height = 0,
        alpha = 0.3, size = 1
    ) +
    # Vertical reference lines at measurement days
    geom_vline(xintercept = c(7, 14, 20, 27), 
               linetype = "dotted", color = "gray70", linewidth = 0.4) +
    scale_color_manual(values = treatment_colors) +
    scale_fill_manual(values = treatment_colors) +
    scale_x_continuous(breaks = c(0, 7, 14, 20, 27, 35)) +
    labs(
        x = "Days from start of experiment",
        y = "Plant height (cm)",
        color = "Treatment",
        fill  = "Treatment",
        title = "Population-level Gompertz growth curves",
        subtitle = "Posterior median with 50% and 90% credible intervals; raw plant heights overlaid"
    ) +
    theme_bayes(base_size = 14) +
    theme(legend.position = "bottom")
```