---
title: "Maize JIP Models"
subtitle: " Maize Models"
author: "Metodi Latinov"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-tools: true
    theme: cosmo
    fig-width: 8
    fig-height: 5
    fig-cap-location: bottom
execute:
  echo: true
  warning: false
  message: false
  cache: false
knitr:
  opts_chunk:
    fig.align: center
---
```{r setup}
cat("Working directory:", getwd(), "\n")
cat("_targets dir exists:", dir.exists("_targets"), "\n")
cat("_targets.R exists:", file.exists("_targets.R"), "\n")
cat("Files here:", paste(head(list.files()), collapse = ", "), "\n")

# Libraries
library(targets)
library(patchwork)
library(tidyverse)
library(stanviz)
library(posterior)

# Load the cleaned data from the pipeline and models
tar_load(model_data_jip)

## Jit Models 
tar_load(jip_model)
tar_load(jip_model_pp)

```

# Jip Models 
Four bounded primaries (gammaRC, phi_Po, psi_Eo, delta_Ro) share one Beta-logit skeleton: outcome ~ Beta(mu*kappa_D, (1-mu)*kappa_D), logit(mu) = eta[treatment,day] + u_pot, non-centered pot offsets, day-specific kappa. RC/CSo (LogNormal, unbounded) is a separate submodel, not yet fit here. PIabs and PItotal are generated quantities, computed per posterior draw from the four primaries via verified algebraic identities — never modeled directly, to avoid double-counting information already in the primaries.

## Jit Model GammaRC
gammaRC — probability a chlorophyll molecule sits in an active reaction center vs. antenna pigment. Feeds Formula 1 as RC/ABS = gammaRC/(1-gammaRC). Expected to be flat: literature treats gammaRC as a conserved structural parameter that barely moves in unstressed plants. Our cell means confirm this: ~0.27-0.29 across all treatment×day cells, tiny spread. Low R² (0.235, logit-scale/McKelvey-Zavoina) is consistent with this — not a fit problem, it's the mathematical signature of "little between-cell variance to explain." First model fit and validated (R-hat 1.00, ESS>14000, 1/12000 divergences).

### Prior Predictive Checks 
Priors are loose and weekly requlizing but they put the prob mass around where it shoud be.Predictive sd is a missaling a bit but is still reasonable 
```{r}
dens_gamma_rc_s <- plot_prior_predictive_dens(
    model = jip_model_pp,
    yrep_var = "rc_rep",
    y = model_data_jip$gammaRC,
    outcome_label = "GammaRC"  
)

gamma_rc_mean <- plot_prior_predictive_stat(
    model = jip_model_pp,
    yrep_var = "rc_rep",
    y = model_data_jip$gammaRC,
    stat = c("mean"),
    outcome_label = "GammaRC" 
)
gamma_rc_sd <- plot_prior_predictive_stat(
    model = jip_model_pp,
    yrep_var = "rc_rep",
    y = model_data_jip$gammaRC,
    stat = c("sd"),
    outcome_label = "GammaRC" 
)
dens_gamma_rc_s / (gamma_rc_mean + gamma_rc_sd) + 
    plot_annotation(
        title = "Prior Predictive Checks for GammaRC Submodel",
        theme = theme(
            plot.title = element_text(face = "bold",hjust = 0.5,size = 15)
        )
    )

```

### Posterior Predictive Checks 
Posterior Predictive checks looks good the prob mass is where it should be still sd is a touch out of the center but again is mostly fine for all intended purposes
```{r}
post_dens_gammarc_s <- plot_ppc_dens(
    model = jip_model,
    y = model_data_jip$gammaRC,
    yrep_var = "rc_rep",
    outcome_label = "Posterior GammaRc"
)
post_stat_gammarc_s <- plot_ppc_stat_grid(
    model = jip_model,
    y = model_data_jip$gammaRC,
    yrep_var = "rc_rep", 
    stats = c("mean","sd")
)
post_dens_gammarc_s / post_stat_gammarc_s + 
    plot_annotation(
        title = "Posterior Predictive Check for GammaRC Submodel",
        theme = theme(
            plot.title = element_text(face = "bold",hjust = 0.5, size = 20)
        )
    )
```

### Posterior Estimands & Construct Coef
```{r}
plot_coef_forest(
    model = jip_model,
    pars  = c(
        "d_rc_emf_ctrl[1]",
        "d_rc_emf_ctrl[2]",
        "d_rc_emf_ctrl[3]"
    ),
    labels = c(
        "d_rc_emf_ctrl[1]" = "Day 13 EMF-Control",
        "d_rc_emf_ctrl[2]" = "Day 20 EMF-Control",
        "d_rc_emf_ctrl[3]" = "Day 27 EMF-Control"
    ) 
)
```

### R2 
```{r}
plot_param_halfeye(
    model = jip_model,
    pars  = "R2_rc_full",
    labels = c("R2_rc_full"= "Posterior R2"),
    null_line = 0.235 
)
```

## Jip Model Phi Po 
phi_Po — max quantum yield of primary photochemistry (classic Fv/Fm family, most-reported/most-stable JIP parameter). Expected flat, like gammaRC. Cell means ~0.72-0.75, no consistent day trend (small day-20 dip in 2/3 treatments, provisionally read as noise — recheck this once Day-20 decomposition for PIabs is run, since this dip is a candidate driver of the PIabs Day-20 signal). R² 0.199, same conserved-parameter interpretation as gammaRC.

## Prior Predictive Checks
Again close picture to the previus submodel Here prior predictive mean and sd are bit off from the center but still for weekly informative prior are reasonable 
```{r}
phi_po_s <- plot_prior_predictive_dens(
    model    = jip_model_pp,
    yrep_var = "po_rep",
    y        = model_data_jip$phi_Po,
    outcome_label = "Prior Predictive Density Check"
)
phi_po_mean_s <- plot_prior_predictive_stat(
    model    = jip_model_pp,
    yrep_var = "po_rep",
    y        = model_data_jip$phi_Po,
    stat     = "mean",
    outcome_label = "Prior Predictive Mean"
)
phi_po_sd_s <- plot_prior_predictive_stat(
    model    = jip_model_pp,
    yrep_var = "po_rep",
    y        = model_data_jip$phi_Po,
    stat     = "sd",
    outcome_label = "Prior Predictive SD"
)

## Combine the Plots 
phi_po_s / (phi_po_mean_s + phi_po_sd_s) +
    plot_annotation(
        title = "Prior Predictive Checks for Phi Po Submodel",
        theme = theme(
            plot.title = element_text(family = "bold",hjust = 0.5, size = 20) 
        )
    )
```

### Posterior Predictive Checks 
Looks resonable simular situation to the Gamma RC model
```{r}
phi_po_p <- plot_ppc_dens(
    model = jip_model,
    y     = model_data_jip$phi_Po,
    yrep_var = "po_rep",
    outcome_label = "Posterior Predictive Distribution"
)
phi_stats <- plot_ppc_stat_grid(
    model = jip_model,
    y     = model_data_jip$phi_Po,
    yrep_var = "po_rep",
    stats    = c("mean", "sd")   
)

## Combine the plots 
phi_po_p / phi_stats + 
    plot_annotation(
        title = "Posterior Predictive Checks",
        theme = theme(
            plot.title = element_text(face = "bold", size = 20, hjust = 0.5
        )
    )  
)
```

### Posterior Estimands & Construct Coef
```{r}
plot_coef_forest(
    model = jip_model,
    pars  = c(
        "d_po_emf_ctrl[1]",
        "d_po_emf_ctrl[2]",
        "d_po_emf_ctrl[3]"
    ),
    labels = c(
        "d_po_emf_ctrl[1]" = "Day 13 EMF-Control",
        "d_po_emf_ctrl[2]" = "Day 20 EMF-Control",
        "d_po_emf_ctrl[3]" = "Day 27 EMF-Control"
    ) 
)
```

### R2 
```{r}
plot_param_halfeye(
    model = jip_model,
    pars  = "R2_po_full",
    labels = c("R2_po_full"= "Posterior R2"),
    null_line = 0.199 
)
```

## Jip Model Psi Eo
psi_Eo — efficiency of electron transfer from QA onward to plastoquinone. Unlike gammaRC/phi_Po, shows a real day trend: rises day 13→20 in all three treatments, then plateaus 20→27 (Control 0.585→0.630→0.632; EMF 0.584→0.612→0.628; Sham 0.601→0.609→0.627). Treatment-consistent shape → real developmental signal, not noise. R² 0.238 — moderate, reflects that day carries real explanatory weight here unlike the two conserved parameters above.

### Prior Predictive Checks
again a bit of miss alighment but still resoable for weak priors
```{r}
psi_eo_s <- plot_prior_predictive_dens(
    model    = jip_model_pp,
    yrep_var = "eo_rep",
    y        = model_data_jip$psi_Eo,
    outcome_label = "Prior Predictive Distribution"
)
psi_eo_mean_s <- plot_prior_predictive_stat(
    model    = jip_model_pp,
    yrep_var = "eo_rep",
    y        = model_data_jip$psi_Eo,
    stat     = "mean",
    outcome_label = "Prior Predictive Mean"
)
psi_eo_sd_s <- plot_prior_predictive_stat(
    model    = jip_model_pp,
    yrep_var = "eo_rep",
    y        = model_data_jip$psi_Eo,
    stat     = "sd",
    outcome_label = "Prior Predictive SD"
)

## Combine the plots 
psi_eo_s / (psi_eo_mean_s + psi_eo_sd_s) +
    plot_annotation(
        title = "Prior Predictive Checks for PSI EO Submodel",
        theme = theme(
            plot.title = element_text(
                size = 20, face = "bold", hjust = 0.5
        )
    ) 
)
```

### Posterior Predictive Check
```{r}
psi_eo_p <- plot_ppc_dens(
    model = jip_model,
    y     = model_data_jip$psi_Eo,
    yrep_var      = "eo_rep",
    outcome_label = "Posterior Predictive Density"
)
psi_eo_stats <- plot_ppc_stat_grid(
    model = jip_model,
    y     = model_data_jip$psi_Eo,
    yrep_var = "eo_rep",
    stats = c("mean", "sd")
)

## Combine the plots
psi_eo_p / psi_eo_stats + 
    plot_annotation(
        title = "Posterior Predictive Checks for PSI EO Submodel",
        theme = theme(
            plot.title = element_text(
                family = "bold", size = 20, hjust = 0.5
            )
    ) 
)
```

### Posterior Estimands & Construct Coef
```{r}
plot_coef_forest(
    model = jip_model,
    pars  = c(
        "d_eo_emf_ctrl[1]",
        "d_eo_emf_ctrl[2]",
        "d_eo_emf_ctrl[3]"
    ),
    labels = c(
        "d_eo_emf_ctrl[1]" = "Day 13 EMF-Control",
        "d_eo_emf_ctrl[2]" = "Day 20 EMF-Control",
        "d_eo_emf_ctrl[3]" = "Day 27 EMF-Control"
    ) 
)
```

### R2 
```{r}
plot_param_halfeye(
    model = jip_model,
    pars  = "R2_eo_full",
    labels = c("R2_eo_full"= "Posterior R2"),
    null_line = 0.238 
)
```

## Jip Model Delta Ro
delta_Ro — efficiency of electron transport to the PSI-side acceptors, the most downstream JIP parameter. Sharp, treatment-consistent drop day 13→20, then plateau 20→27 (Control 0.514→0.346; EMF 0.529→0.355; Sham 0.516→0.374) — this is what produces the marginal-bimodal look in the raw prior-predictive check (two piles: day-13 values ~0.5, day-20/27 values ~0.35-0.37). Resolved structurally by conditioning on day — confirmed each treatment×day cell is unimodal (see posterior predictive, which is clean). Mechanically drives a large swing in PItotal's odds term specifically at day 13. Negatively correlated with the other three primaries (residual r ≈ -0.25/-0.09/-0.46), so PItotal's uncertainty correction from the eventual joint model will pull the opposite direction from PIabs's.

### Prior Predictive 
Possible the worst because the distribution is bimodal for some reason (day) .... again still works
```{r}
delta_ro_s <- plot_prior_predictive_dens(
    model = jip_model_pp,
    yrep_var = "ro_rep",
    y        = model_data_jip$delta_Ro,
    outcome_label = "Prior Predictive Distribution"
)
delta_ro_mean_s <- plot_prior_predictive_stat(
    model = jip_model_pp,
    yrep_var = "ro_rep",
    y        = model_data_jip$delta_Ro,
    stat     = "mean",
    outcome_label = "Prior Predictive Mean"
)
delta_ro_sd_s <- plot_prior_predictive_stat(
    model = jip_model_pp,
    yrep_var = "ro_rep",
    y        = model_data_jip$delta_Ro,
    stat     = "sd",
    outcome_label = "Prior Predictive SD"
)

## Combine the plots
delta_ro_s / (delta_ro_mean_s + delta_ro_sd_s)+
    plot_annotation(
        title = "Prior Predictive Checks for Delta Ro Submodel",
        theme = theme(
            plot.title = element_text(face = "bold", size = 20, hjust = 0.5
        )
    )
)
```

### Posterior Predictive Checks 
```{r}
delta_p <- plot_ppc_dens(
    model = jip_model,
    y     = model_data_jip$delta_Ro,
    yrep_var      = "ro_rep",
    outcome_label = "Posterior Predictive Density"
)
delta_ro_stats <- plot_ppc_stat_grid(
    model = jip_model,
    y     = model_data_jip$delta_Ro,
    yrep_var = "ro_rep",
    stats = c("mean", "sd")
)

## Combine the plots
delta_p / delta_ro_stats + 
    plot_annotation(
        title = "Posterior Predictive Checks for Delta Ro Submodel",
        theme = theme(
            plot.title = element_text(
                family = "bold", size = 20, hjust = 0.5
            )
    ) 
)
```

### Posterior Estimands & Construct Coef
```{r}
plot_coef_forest(
    model = jip_model,
    pars  = c(
        "d_ro_emf_ctrl[1]",
        "d_ro_emf_ctrl[2]",
        "d_ro_emf_ctrl[3]"
    ),
    labels = c(
        "d_ro_emf_ctrl[1]" = "Day 13 EMF-Control",
        "d_ro_emf_ctrl[2]" = "Day 20 EMF-Control",
        "d_ro_emf_ctrl[3]" = "Day 27 EMF-Control"
    ) 
)
```

# Estimands Derived from the Combined JIP Models 

## PIabs
PIabs — absorption-based performance index: RC/ABS × [phi_Po odds] × [psi_Eo odds], where RC/ABS = gammaRC/(1-gammaRC). Composite of reaction-center density × two downstream efficiencies. Generated quantity, computed per posterior draw from the three constituent primaries — identity verified to machine precision against the raw data. Forest plot shows Day 20 EMF-Control and Sham-Control both negative (medians ≈ -0.11 to -0.13), while Day 20 EMF-Sham sits at ~0 — pattern points to a room/non-Control effect at day 20 rather than an EMF-specific one, echoing the height Sham-deficit story. Candidate driver: phi_Po's day-20 dip (see above) — needs the primary-by-primary Day 20 decomposition to confirm.

### Posterior Estimates & Posterior Contracts per Treatment + Deconstruction
```{r}
plot_coef_forest(
    model = jip_model,
    pars   = c(
        "logPIabs_emf_ctrl[1]","logPIabs_emf_ctrl[2]","logPIabs_emf_ctrl[3]"
        ),
    labels = c(
        "logPIabs_emf_ctrl[1]" = "Day 13 EMF-Control",
        "logPIabs_emf_ctrl[2]" = "Day 20 EMF-Control",
        "logPIabs_emf_ctrl[3]" = "Day 27 EMF-Control"
    ),
    sort = TRUE
)
```

## PItotal
PItotal — total performance index: PIabs × [delta_Ro odds]. Extends PIabs one step further through the PSI acceptor side. Generated quantity, same machine-precision-verified identity. Because delta_Ro correlates negatively with the other three primaries (unlike PIabs's all-positive-correlation components), expect PItotal's joint-model uncertainty correction to move in the opposite direction from PIabs's once the capstone model is built — this is a reportable result in its own right, not just a footnote.

### Posterior Estimates & Posterior Contracts per Treatment + Deconstruction
```{r}
plot_coef_forest(
    model = jip_model,
    pars  = c(
        "logPItot_emf_ctrl[1]",
        "logPItot_emf_ctrl[2]",
        "logPItot_emf_ctrl[3]"
    ),
    labels = c(
        "logPItot_emf_ctrl[1]" = "Day 13 EMF-Control",
        "logPItot_emf_ctrl[2]" = "Day 20 EMF-Control",
        "logPItot_emf_ctrl[3]" = "Day 27 EMF-Control"
    ),
    sort = TRUE
)
```