---
title: "Maize Height Models Checks"
subtitle: "Comparison Between Different Maize Models"
author: "Metodi Latinov"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-tools: true
    theme: cosmo
    fig-width: 8
    fig-height: 5
    fig-cap-location: bottom
execute:
  echo: true
  warning: false
  message: false
  cache: false
knitr:
  opts_chunk:
    fig.align: center
---

```{r setup}
cat("Working directory:", getwd(), "\n")
cat("_targets dir exists:", dir.exists("_targets"), "\n")
cat("_targets.R exists:", file.exists("_targets.R"), "\n")
cat("Files here:", paste(head(list.files()), collapse = ", "), "\n")

# Libraries
library(targets)
library(patchwork)
library(tidyverse)
library(stanviz)

# Load the cleaned data from the pipeline and models
tar_load(model_data_maize_heigh)

## M1  Plant height ~ Treatment only
tar_load(m1_maize_model)
tar_load(m1_maize_model_pp)

## M2 Plant height ~  Treatment × Day
tar_load(m2_maize_model)
tar_load(m2_maize_model_pp)

## M3 Plant height ~ Pot x Day
tar_load(m3_maize_model)
tar_load(m3_maize_model_pp)

## M4 
tar_load(m4_maize_model)
tar_load(m4_maize_model_pp)

```

# M1  — Treatment only

`plant_height ~ treatment`, with `plant_height ~ LogNormal(α[treatment], σ)` and 
weakly-informative priors `α ~ Normal(3, 1)`, `σ ~ Exponential(1)`. 

This is the deliberately naive baseline (M1.1 in the project plan): only treatment, no 
day, no pot. We *expect* it to be misspecified, because we know the data is structured 
by day. The point of fitting it is to (1) validate the indexing, priors, and PPC 
machinery on the simplest possible model, and (2) make the misspecification visible 
so the next rung (adding day) is empirically motivated rather than assumed.

## Model Diagnostics

### MCMC Sampler Diagnostics

Trace ranks should look uniform across chains — no chain occupying systematically 
high or low ranks. With a model this simple the posterior geometry is trivial and 
mixing is essentially perfect.

```{r}
plot_rank(model = m1_maize_model)
```

`cmdstan_diagnose()` reported no problems: no divergences, treedepth and E-BFMI 
satisfactory, R-hat and ESS within thresholds for every parameter. The ESS plots 
below appear to alarm visually but the numerical thresholds are not violated — 
bulk-ESS and tail-ESS for every model parameter are in the thousands (out of 4000 
post-warmup draws), and every R-hat rounds to 1.00. The colour-banding crosses 0.5 
at the dramatic-looking boundary but the meaningful danger threshold is 0.1.

```{r}
m1_r_hat    <- plot_rhat(model = m1_maize_model)
m1_ess_tail <- plot_ess(model = m1_maize_model, type = "tail")
m1_ess_bulk <- plot_ess(model = m1_maize_model, type = "bulk")

(m1_ess_tail + m1_ess_bulk) / m1_r_hat +
    plot_annotation(
        title = "Effective Sample Size and R-hats — Simple Maize Model",
        theme = theme(plot.title = element_text(size = 20))
    )
```

**Sampler verdict: clean.** Any failure we see below is a model-specification 
issue, not an MCMC issue.

### Prior Predictive Checks

Before looking at the posterior, we check that the priors alone produce data 
that looks vaguely like maize. The prior-predictive mean sits around 24 cm and 
the SD around 17 cm; the observed values fall well inside both prior-predictive 
distributions. The bottom panel overlays prior-implied datasets against the 
observed density: the observed line sits comfortably inside the cloud of 
prior-predictive curves. Priors are loose but not absurd — they don't rule out 
the data before we even see it.

```{r}
m1_pr_dens <- plot_prior_predictive_dens(
    model         = m1_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
)

m1_pr_mean <- plot_prior_predictive_stat(
    model         = m1_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "mean"
)

m1_pr_sd <- plot_prior_predictive_stat(
    model         = m1_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "sd"
)

(m1_pr_mean + m1_pr_sd) / m1_pr_dens +
    plot_annotation(
        title = "Prior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

### Posterior Predictive Checks

This is where M1.1 fails — exactly as the modelling ladder predicted. Three 
independent symptoms, all pointing at the same cause:

1. **Density (bottom panel).** The observed distribution is clearly **bimodal**: 
   two humps, corresponding to small early-day plants and tall late-day plants 
   piled into the same marginal. Every posterior replicate is a smooth 
   single-mode lognormal. M1.1 has no day parameter, so it cannot reproduce 
   the bimodal structure in principle.
2. **SD test statistic.** The posterior predictive SD is centred around 17 cm 
   while the observed SD is 13 cm — the model over-predicts spread because 
   `σ` is being forced wide enough to absorb day-7-to-day-27 growth as if it 
   were within-group noise.
3. **Posterior predictive tail (not shown — visible by inspecting `range(h_rep)`).** 
   With `σ ≈ 0.62` on the log scale, the right tail of the lognormal is 
   heavy enough to produce occasional draws orders of magnitude larger than 
   any real maize plant. This is the model being internally consistent with a 
   parameter value that was itself forced wrong by the missing day structure.

All three symptoms have the same root cause: M1.1 absorbs between-day 
variation into within-group variation. The fix is structural, not 
likelihood-side — adding day to the mean function (M1.2) is predicted to 
collapse `σ`, eliminate the heavy tail, and at least partially recover the 
bimodality. That prediction is the validation criterion for M1.2.

```{r}
m1_ppc_dens <- plot_ppc_dens(
    model         = m1_maize_model,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
)

m1_ppc_mean <- plot_ppc_stat(
    model    = m1_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "mean"
) + coord_cartesian(xlim = c(15, 35))

m1_ppc_sd <- plot_ppc_stat(
    model    = m1_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m1_ppc_mean + m1_ppc_sd) / m1_ppc_dens +
    plot_annotation(
        title = "Posterior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

# M2 — Treatment × Day

## Model Diagnostics

### MCMC Sampler Diagnostics

Trace ranks should look uniform across chains — no chain occupying systematically 
high or low ranks. With a model this simple the posterior geometry is trivial and 
mixing is essentially perfect.

```{r}
plot_rank(model = m2_maize_model)
```

`cmdstan_diagnose()` reported no problems: no divergences, treedepth and E-BFMI 
satisfactory, R-hat and ESS within thresholds for every parameter. The ESS plots 
below appear to alarm visually but the numerical thresholds are not violated — 
bulk-ESS and tail-ESS for every model parameter are in the thousands (out of 4000 
post-warmup draws), and every R-hat rounds to 1.00. The colour-banding crosses 0.5 
at the dramatic-looking boundary but the meaningful danger threshold is 0.1.

```{r}
m2_r_hat    <- plot_rhat(model = m2_maize_model)
m2_ess_tail <- plot_ess(model = m2_maize_model, type = "tail")
m2_ess_bulk <- plot_ess(model = m2_maize_model, type = "bulk")

(m2_ess_tail + m2_ess_bulk) / m2_r_hat +
    plot_annotation(
        title = "Effective Sample Size and R-hats — Simple Maize Model",
        theme = theme(plot.title = element_text(size = 20))
    )
```

**Sampler verdict: clean.** Any failure we see below is a model-specification 
issue, not an MCMC issue.

### Prior Predictive Checks

```{r}
m2_pr_dens <- plot_prior_predictive_dens(
    model         = m2_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
)

m2_pr_mean <- plot_prior_predictive_stat(
    model         = m2_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "mean"
)

m2_pr_sd <- plot_prior_predictive_stat(
    model         = m2_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "sd"
)

(m2_pr_mean + m2_pr_sd) / m2_pr_dens +
    plot_annotation(
        title = "Prior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

### Posterior Predictive Checks

This is where M1.1 fails — exactly as the modelling ladder predicted. Three 
independent symptoms, all pointing at the same cause:

1. **Density (bottom panel).** The observed distribution is clearly **bimodal**: 
   two humps, corresponding to small early-day plants and tall late-day plants 
   piled into the same marginal. Every posterior replicate is a smooth 
   single-mode lognormal. M1.1 has no day parameter, so it cannot reproduce 
   the bimodal structure in principle.
2. **SD test statistic.** The posterior predictive SD is centred around 17 cm 
   while the observed SD is 13 cm — the model over-predicts spread because 
   `σ` is being forced wide enough to absorb day-7-to-day-27 growth as if it 
   were within-group noise.
3. **Posterior predictive tail (not shown — visible by inspecting `range(h_rep)`).** 
   With `σ ≈ 0.62` on the log scale, the right tail of the lognormal is 
   heavy enough to produce occasional draws orders of magnitude larger than 
   any real maize plant. This is the model being internally consistent with a 
   parameter value that was itself forced wrong by the missing day structure.

All three symptoms have the same root cause: M1.1 absorbs between-day 
variation into within-group variation. The fix is structural, not 
likelihood-side — adding day to the mean function (M1.2) is predicted to 
collapse `σ`, eliminate the heavy tail, and at least partially recover the 
bimodality. That prediction is the validation criterion for M1.2.

```{r}
m2_ppc_dens <- plot_ppc_dens(
    model         = m2_maize_model,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
)

m2_ppc_mean <- plot_ppc_stat(
    model    = m2_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "mean"
) + coord_cartesian(xlim = c(15, 35))

m2_ppc_sd <- plot_ppc_stat(
    model    = m2_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m2_ppc_mean + m2_ppc_sd) / m2_ppc_dens +
    plot_annotation(
        title = "Posterior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

# M3 — Pot × Day

## Model Diagnostics

### MCMC Sampler Diagnostics

Trace ranks should look uniform across chains — no chain occupying systematically 
high or low ranks. With a model this simple the posterior geometry is trivial and 
mixing is essentially perfect.

```{r}
plot_rank(model = m3_maize_model)
```

`cmdstan_diagnose()` reported no problems: no divergences, treedepth and E-BFMI 
satisfactory, R-hat and ESS within thresholds for every parameter. The ESS plots 
below appear to alarm visually but the numerical thresholds are not violated — 
bulk-ESS and tail-ESS for every model parameter are in the thousands (out of 4000 
post-warmup draws), and every R-hat rounds to 1.00. The colour-banding crosses 0.5 
at the dramatic-looking boundary but the meaningful danger threshold is 0.1.

```{r}
m3_r_hat    <- plot_rhat(model = m3_maize_model)
m3_ess_tail <- plot_ess(model = m3_maize_model, type = "tail")
m3_ess_bulk <- plot_ess(model = m3_maize_model, type = "bulk")

(m3_ess_tail + m3_ess_bulk) / m3_r_hat +
    plot_annotation(
        title = "Effective Sample Size and R-hats — M3 Maize Model",
        theme = theme(plot.title = element_text(size = 20))
    )
```

**Sampler verdict: clean.** Any failure we see below is a model-specification 
issue, not an MCMC issue.

### Prior Predictive Checks

```{r}
m3_pr_dens <- plot_prior_predictive_dens(
    model         = m3_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
)

m3_pr_mean <- plot_prior_predictive_stat(
    model         = m3_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "mean"
)

m3_pr_sd <- plot_prior_predictive_stat(
    model         = m3_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "sd"
)

(m3_pr_mean + m3_pr_sd) / m3_pr_dens +
    plot_annotation(
        title = "Prior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

### Posterior Predictive Checks

This is where M1.1 fails — exactly as the modelling ladder predicted. Three 
independent symptoms, all pointing at the same cause:

1. **Density (bottom panel).** The observed distribution is clearly **bimodal**: 
   two humps, corresponding to small early-day plants and tall late-day plants 
   piled into the same marginal. Every posterior replicate is a smooth 
   single-mode lognormal. M1.1 has no day parameter, so it cannot reproduce 
   the bimodal structure in principle.
2. **SD test statistic.** The posterior predictive SD is centred around 17 cm 
   while the observed SD is 13 cm — the model over-predicts spread because 
   `σ` is being forced wide enough to absorb day-7-to-day-27 growth as if it 
   were within-group noise.
3. **Posterior predictive tail (not shown — visible by inspecting `range(h_rep)`).** 
   With `σ ≈ 0.62` on the log scale, the right tail of the lognormal is 
   heavy enough to produce occasional draws orders of magnitude larger than 
   any real maize plant. This is the model being internally consistent with a 
   parameter value that was itself forced wrong by the missing day structure.

All three symptoms have the same root cause: M1.1 absorbs between-day 
variation into within-group variation. The fix is structural, not 
likelihood-side — adding day to the mean function (M1.2) is predicted to 
collapse `σ`, eliminate the heavy tail, and at least partially recover the 
bimodality. That prediction is the validation criterion for M1.2.

```{r}
m3_ppc_dens <- plot_ppc_dens(
    model         = m3_maize_model,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
)

m3_ppc_mean <- plot_ppc_stat(
    model    = m3_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "mean"
) + coord_cartesian(xlim = c(15, 35))

m3_ppc_sd <- plot_ppc_stat(
    model    = m3_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m3_ppc_mean + m3_ppc_sd) / m3_ppc_dens +
    plot_annotation(
        title = "Posterior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

```{r}
pot_medians <- m3_maize_model$summary("alpha") |>
  mutate(
    # parse out pot and day indices from the variable name "alpha[p,d]"
    pot_id = as.integer(stringr::str_extract(variable, "(?<=\\[)\\d+")),
    day    = as.integer(stringr::str_extract(variable, "(?<=,)\\d+(?=\\])")),
    treatment = case_when(
      pot_id <= 6  ~ "Control",
      pot_id <= 10 ~ "Sham",
      TRUE         ~ "EMF"
    )
  )

ggplot(pot_medians, aes(x = factor(day, labels = c("7","14","20","27")), 
                         y = median, color = treatment)) +
  geom_point(position = position_jitter(width = 0.15, height = 0), size = 2.5, alpha = 0.8) +
  geom_linerange(aes(ymin = q5, ymax = q95), 
                 position = position_jitter(width = 0.15, height = 0), alpha = 0.5) +
  labs(x = "Day", y = "log-median height (cm)",
       title = "Per-pot posterior medians by treatment and day",
       subtitle = "Each point is one pot; vertical bars are 90% CI") +
  theme_minimal()
```

# M4 Gompertz per treatment 

## Model Diagnostics

### MCMC Sampler Diagnostics

Trace ranks should look uniform across chains — no chain occupying systematically 
high or low ranks. With a model this simple the posterior geometry is trivial and 
mixing is essentially perfect.

```{r}
plot_rank(model = m4_maize_model)
```

`cmdstan_diagnose()` reported no problems: no divergences, treedepth and E-BFMI 
satisfactory, R-hat and ESS within thresholds for every parameter. The ESS plots 
below appear to alarm visually but the numerical thresholds are not violated — 
bulk-ESS and tail-ESS for every model parameter are in the thousands (out of 4000 
post-warmup draws), and every R-hat rounds to 1.00. The colour-banding crosses 0.5 
at the dramatic-looking boundary but the meaningful danger threshold is 0.1.

```{r}
m4_r_hat    <- plot_rhat(model = m4_maize_model)
m4_ess_tail <- plot_ess(model = m4_maize_model, type = "tail")
m4_ess_bulk <- plot_ess(model = m4_maize_model, type = "bulk")

(m4_ess_tail + m4_ess_bulk) / m4_r_hat +
    plot_annotation(
        title = "Effective Sample Size and R-hats — M4 Maize Model",
        theme = theme(plot.title = element_text(size = 20))
    )
```

**Sampler verdict: clean.** Any failure we see below is a model-specification 
issue, not an MCMC issue.

### Prior Predictive Checks

```{r}
m4_pr_dens <- plot_prior_predictive_dens(
    model         = m4_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
) + coord_cartesian(xlim = c(0, 80))

m4_pr_mean <- plot_prior_predictive_stat(
    model         = m4_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "mean"
) + coord_cartesian(xlim = c(15, 35))

m4_pr_sd <- plot_prior_predictive_stat(
    model         = m4_maize_model_pp,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height",
    stat          = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m4_pr_mean + m4_pr_sd) / m4_pr_dens +
    plot_annotation(
        title = "Prior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```

### Posterior Predictive Checks

This is where M1.1 fails — exactly as the modelling ladder predicted. Three 
independent symptoms, all pointing at the same cause:

1. **Density (bottom panel).** The observed distribution is clearly **bimodal**: 
   two humps, corresponding to small early-day plants and tall late-day plants 
   piled into the same marginal. Every posterior replicate is a smooth 
   single-mode lognormal. M1.1 has no day parameter, so it cannot reproduce 
   the bimodal structure in principle.
2. **SD test statistic.** The posterior predictive SD is centred around 17 cm 
   while the observed SD is 13 cm — the model over-predicts spread because 
   `σ` is being forced wide enough to absorb day-7-to-day-27 growth as if it 
   were within-group noise.
3. **Posterior predictive tail (not shown — visible by inspecting `range(h_rep)`).** 
   With `σ ≈ 0.62` on the log scale, the right tail of the lognormal is 
   heavy enough to produce occasional draws orders of magnitude larger than 
   any real maize plant. This is the model being internally consistent with a 
   parameter value that was itself forced wrong by the missing day structure.

All three symptoms have the same root cause: M1.1 absorbs between-day 
variation into within-group variation. The fix is structural, not 
likelihood-side — adding day to the mean function (M1.2) is predicted to 
collapse `σ`, eliminate the heavy tail, and at least partially recover the 
bimodality. That prediction is the validation criterion for M1.2.

```{r}
m4_ppc_dens <- plot_ppc_dens(
    model         = m4_maize_model,
    yrep_var      = "h_rep",
    y             = model_data_maize_heigh$plant_height,
    outcome_label = "Maize Height"
) + coord_cartesian(xlim = c(0, 80))

m4_ppc_mean <- plot_ppc_stat(
    model    = m4_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "mean"
) + coord_cartesian(xlim = c(15, 35))

m4_ppc_sd <- plot_ppc_stat(
    model    = m4_maize_model,
    yrep_var = "h_rep",
    y        = model_data_maize_heigh$plant_height,
    stat     = "sd"
) + coord_cartesian(xlim = c(10, 25))

(m4_ppc_mean + m4_ppc_sd) / m4_ppc_dens +
    plot_annotation(
        title = "Posterior Predictive Density and Summary Statistics",
        theme = theme(plot.title = element_text(size = 20))
    )
```
