---
title: "Maize Biochemistry-Biomass Models"
subtitle: "Mediation Maize Models"
author: "Metodi Latinov"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-tools: true
    theme: cosmo
    fig-width: 8
    fig-height: 5
    fig-cap-location: bottom
execute:
  echo: true
  warning: false
  message: false
  cache: false
knitr:
  opts_chunk:
    fig.align: center
---

```{r setup}
cat("Working directory:", getwd(), "\n")
cat("_targets dir exists:", dir.exists("_targets"), "\n")
cat("_targets.R exists:", file.exists("_targets.R"), "\n")
cat("Files here:", paste(head(list.files()), collapse = ", "), "\n")

#### Libraries ####
library(targets)
library(stanviz)
library(patchwork)
library(tidyverse)

#### Load the Targets 
tar_load(data_bb)
tar_load(model_bb_base)
tar_load(model_bb_pp_base)

```

# Prior Predictive Checks Primary Models 
Prior Predictive of the Primary Models looks to approximates the sample well worth noted in the document the long tails that were clipped for the visualization.
```{r}
sod_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[1,]",
  y        =  data_bb$data$sod_dw,
  outcome_label = "SOD, U/mg DW"
)
cat_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[2,]",
  y        = data_bb$data$cat_dw,
  outcome_label = "CAT, µmol H2O2/min/mg DW"
) + xlim(c(0, 10))

sugar_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[3,]",
  y        = data_bb$data$reduced_sugar_dw,
  outcome_label = "Reducing Sugars, mg Glucose/g DW"   
)+ xlim(c(0, 150))

trolox_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[4,]",
  y        = data_bb$data$trolox_umol_dw,
  outcome_label = "Trolox Equivalent, µmol/g DW" 
)

## Combine the plots ##
(sod_dw_ppc + cat_dw_ppc) / (sugar_dw_ppc + trolox_dw_ppc) +
  plot_annotation(
    title = "Prior Predictive Distributions of Primary Models",
    theme = theme(
      plot.title = element_text(face = "italic", size = 20,hjust = 0.5)
    )
  )
```

# Posterior Predictive Checks Primary Models 
Prior Distributions Long Tail problems is fixed in the posterior No clipping required.
The Posterior Accurately approximate the sample 
```{r}
sod_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[1,]",
  y        =  data_bb$data$sod_dw,
  outcome_label = "SOD, U/mg DW"
)
cat_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[2,]",
  y        = data_bb$data$cat_dw,
  outcome_label = "CAT, µmol H2O2/min/mg DW"
) 
sugar_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[3,]",
  y        = data_bb$data$reduced_sugar_dw,
  outcome_label = "Reducing Sugars, mg Glucose/g DW"   
)
trolox_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[4,]",
  y        = data_bb$data$trolox_umol_dw,
  outcome_label = "Trolox Equivalent, µmol/g DW" 
)

## Combine the plots ##
(sod_dw_ppc + cat_dw_ppc) / (sugar_dw_ppc + trolox_dw_ppc) +
  plot_annotation(
    title = "Posterior Predictive ECDF of Primary Models",
    theme = theme(
      plot.title = element_text(face = "italic", size = 20,hjust = 0.5)
    )
  )
```

```{r}
sod_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[1,]",
  y        =  data_bb$data$sod_dw,
  outcome_label = "SOD, U/mg DW"
)
cat_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[2,]",
  y        = data_bb$data$cat_dw,
  outcome_label = "CAT, µmol H2O2/min/mg DW"
) 
sugar_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[3,]",
  y        = data_bb$data$reduced_sugar_dw,
  outcome_label = "Reducing Sugars, mg Glucose/g DW"   
)
trolox_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[4,]",
  y        = data_bb$data$trolox_umol_dw,
  outcome_label = "Trolox Equivalent, µmol/g DW" 
)

## Combine the plots ##
(sod_dw_ppc + cat_dw_ppc) / (sugar_dw_ppc + trolox_dw_ppc) +
  plot_annotation(
    title = "Posterior Predictive Distributions of Primary Models",
    theme = theme(
      plot.title = element_text(face = "italic", size = 20,hjust = 0.5)
    )
  )
```

# Prior Predictive Checks for the Rest of the Models 
```{r}
h2o2_dw_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[5,]",
  y        =  data_bb$data$h2_o2_dw,
  outcome_label = "H2O2, nmol/g DW"
)+ xlim(c(0, 300))
mda_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[6,]",
  y        = data_bb$data$mda_dw,
  outcome_label = "MDA, nmol/g DW"
)+xlim(0, 1000)

water_dw_ppc <- stanviz::plot_prior_predictive_dens(
  model    = model_bb_pp_base$sample,
  yrep_var = "samples_rep[7,]",
  y        = data_bb$data$water_content_dw,
  outcome_label = "Leaf Water Content Relative to DW"   
)+ xlim(0, 20)

## Combine the plots ##
(h2o2_dw_dw_ppc + mda_dw_ppc) / water_dw_ppc +
  plot_annotation(
    title = "Prior Predictive Distributions Dependent Models ",
    theme = theme(
      plot.title = element_text(face = "italic", size = 20,hjust = 0.5)
    )
  )
```

# Posterior Predictive Distribution of the Dependent Models
```{r}
h2o2_dw_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[5,]",
  y        =  data_bb$data$h2_o2_dw,
  outcome_label = "H2O2, nmol/g DW"
)
mda_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[6,]",
  y        = data_bb$data$mda_dw,
  outcome_label = "MDA, nmol/g DW"
)

water_dw_ppc <- stanviz::plot_ppc_dens(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[7,]",
  y        = data_bb$data$water_content_dw,
  outcome_label = "Leaf Water Content Relative to DW"   
)

## Combine the plots ##
(h2o2_dw_dw_ppc + mda_dw_ppc) / water_dw_ppc +
  plot_annotation(
    title = "Posterior Predictive Distributions Dependent Models ",
    theme = theme(
      plot.title = element_text(face = "italic", size = 20,hjust = 0.5)
    )
  )
```

```{r}
h2o2_dw_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[5,]",
  y        =  data_bb$data$h2_o2_dw,
  outcome_label = "H2O2, nmol/g DW"
)
mda_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[6,]",
  y        = data_bb$data$mda_dw,
  outcome_label = "MDA, nmol/g DW"
)

water_dw_ppc <- stanviz::plot_ppc_ecdf(
  model    = model_bb_base$sample,
  yrep_var = "samples_rep[7,]",
  y        = data_bb$data$water_content_dw,
  outcome_label = "Leaf Water Content Relative to DW"   
)

## Combine the plots ##
(h2o2_dw_dw_ppc + mda_dw_ppc) / water_dw_ppc +
  plot_annotation(
    title = "Posterior Predictive ECDF Dependent Models ",
    theme = theme(
      plot.title = element_text(face = "italic", size = 20,hjust = 0.5)
    )
  )
```

# Treatment contrasts

## SOD
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[1,1]", "d_emf_ctrl[1,2]", "d_emf_ctrl[1,3]"),
  labels = c(
    "d_emf_ctrl[1,1]" = "EMF-Control 14",
    "d_emf_ctrl[1,2]" = "EMF-Control 21",
    "d_emf_ctrl[1,3]" = "EMF-Control 28"
  )
)
```

## CAT 
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[2,1]", "d_emf_ctrl[2,2]", "d_emf_ctrl[2,3]"),
  labels = c(
    "d_emf_ctrl[2,1]" = "EMF-Control 14",
    "d_emf_ctrl[2,2]" = "EMF-Control 21",
    "d_emf_ctrl[2,3]" = "EMF-Control 28"
  )
)
```

## TEAC 
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[4,1]", "d_emf_ctrl[4,2]", "d_emf_ctrl[4,3]"),
  labels = c(
    "d_emf_ctrl[4,1]" = "EMF-Control 14",
    "d_emf_ctrl[4,2]" = "EMF-Control 21",
    "d_emf_ctrl[4,3]" = "EMF-Control 28"
  )
)
```


## Sugar
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[3,1]", "d_emf_ctrl[3,2]", "d_emf_ctrl[3,3]"),
  labels = c(
    "d_emf_ctrl[3,1]" = "EMF-Control 14",
    "d_emf_ctrl[3,2]" = "EMF-Control 21",
    "d_emf_ctrl[3,3]" = "EMF-Control 28"
  )
)
```

## H2o2
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[5,1]", "d_emf_ctrl[5,2]", "d_emf_ctrl[5,3]"),
  labels = c(
    "d_emf_ctrl[5,1]" = "EMF-Control 14",
    "d_emf_ctrl[5,2]" = "EMF-Control 21",
    "d_emf_ctrl[5,3]" = "EMF-Control 28"
  )
)
```

## MDA
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[6,1]", "d_emf_ctrl[6,2]", "d_emf_ctrl[6,3]"),
  labels = c(
    "d_emf_ctrl[6,1]" = "EMF-Control 14",
    "d_emf_ctrl[6,2]" = "EMF-Control 21",
    "d_emf_ctrl[6,3]" = "EMF-Control 28"
  )
)
```

## Water
```{r}
stanviz::plot_coef_forest(
  model = model_bb_base$sample,
  pars  = c("d_emf_ctrl[7,1]", "d_emf_ctrl[7,2]", "d_emf_ctrl[7,3]"),
  labels = c(
    "d_emf_ctrl[7,1]" = "EMF-Control 14",
    "d_emf_ctrl[7,2]" = "EMF-Control 21",
    "d_emf_ctrl[7,3]" = "EMF-Control 28"
  )
)
```

# Path coefficients (betas)
```{r}
stanviz::plot_coef_forest(
  model  = model_bb_base$sample,
  pars   = c("beta_sod_dw", "beta_cat_dw", "beta_trolox_dw", "beta_h2o2_dw", "beta_sugar_dw", "beta_mda_dw"),
  labels = c(
    "beta_sod_dw"    = "SOD-H202", 
    "beta_cat_dw"    = "CAT-H202",
    "beta_trolox_dw" = "TEAC-H202",  
    "beta_h2o2_dw"   = "H202-MDA",
    "beta_sugar_dw"  = "SUGAR-Water",
    "beta_mda_dw"    = "MDA-Water"
  )
)
```

# Conditional Effects 

## SOD-H202
```{r}
stanviz::plot_effect_curve(
  model         = model_bb_base$sample,
  effect_var    = "effect_curves[1,]",
  x             = model_bb_base$stan_data$sod_z_grid,                    
  x_obs         = (log(data_bb$data$sod_dw) - mean(log(data_bb$data$sod_dw))) / sd(log(data_bb$data$sod_dw)),
  y_obs         = log(data_bb$data$h2_o2_dw),
  outcome_label = "log H2O2, nmol/g DW",
  x_label       = "SOD, U/mg DW (standardized)"
)
```

## CAT-H2O2
```{r}
stanviz::plot_effect_curve(
  model         = model_bb_base$sample,
  effect_var    = "effect_curves[2,]",
  x             = model_bb_base$stan_data$cat_z_grid,                    
  x_obs         = (log(data_bb$data$cat_dw) - mean(log(data_bb$data$cat_dw))) / sd(log(data_bb$data$cat_dw)),
  y_obs         = log(data_bb$data$h2_o2_dw),
  outcome_label = "log H2O2, nmol/g DW",
  x_label       = "CAT, µmol H2O2/min/mg DW (standardized)"
)
```

## H2O2-MDA
```{r}
stanviz::plot_effect_curve(
  model         = model_bb_base$sample,
  effect_var    = "effect_curves[3,]",
  x             = model_bb_base$stan_data$h202_z_grid,                    
  x_obs         = (log(data_bb$data$h2_o2_dw) - mean(log(data_bb$data$h2_o2_dw))) / sd(log(data_bb$data$h2_o2_dw)),
  y_obs         = log(data_bb$data$mda_dw),
  outcome_label = "log MDA, nmol/g DW",
  x_label       = "H2O2, nmol/g DW (standardized)"
)
```

# Mediation decomposition

## EMF vs Control, day 14
```{r}
plot_indirect_effects(
  model_bb_base$sample,
  nie_vars = c(
    "SOD"  = "TNIE_sod[3,1,1]",
    "CAT"  = "TNIE_cat[3,1,1]",
    "TEAC" = "TNIE_teac[3,1,1]"
  ),
  effect_units = "log H2O2",
  sort = TRUE
)

plot_indirect_effects(
  model_bb_base$sample,
  nie_vars = c(
    "MDA"    = "TNIE_mda_water[3,1,1]",
    "Sugars" = "TNIE_sugar_water[3,1,1]"
  ),
  effect_units = "log water content",
  sort = TRUE
)
```

## EMF vs Control, day 21
```{r}
plot_indirect_effects(
  model_bb_base$sample,
  nie_vars = c(
    "SOD"  = "TNIE_sod[3,1,2]",
    "CAT"  = "TNIE_cat[3,1,2]",
    "TEAC" = "TNIE_teac[3,1,2]"
  ),
  effect_units = "log H2O2",
  sort = TRUE
)

plot_indirect_effects(
  model_bb_base$sample,
  nie_vars = c(
    "MDA"    = "TNIE_mda_water[3,1,2]",
    "Sugars" = "TNIE_sugar_water[3,1,2]"
  ),
  effect_units = "log water content",
  sort = TRUE
)
```

## EMF vs Control, day 28
```{r}
plot_indirect_effects(
  model_bb_base$sample,
  nie_vars = c(
    "SOD"  = "TNIE_sod[3,1,3]",
    "CAT"  = "TNIE_cat[3,1,3]",
    "TEAC" = "TNIE_teac[3,1,3]"
  ),
  effect_units = "log H2O2",
  sort = TRUE
)

plot_indirect_effects(
  model_bb_base$sample,
  nie_vars = c(
    "MDA"    = "TNIE_mda_water[3,1,3]",
    "Sugars" = "TNIE_sugar_water[3,1,3]"
  ),
  effect_units = "log water content",
  sort = TRUE
)
```

# TE
```{r}
plot_effect_decomposition(
  model_bb_base$sample,
  direct_var   = "PNDE_h2o2[3,1,1]",
  indirect_var = "TNIE_h2o2[3,1,1]",
  total_var    = "TE_h2o2[3,1,1]",
  effect_units = "log H2O2"
)

plot_effect_decomposition(
  model_bb_base$sample,
  direct_var   = "PNDE_mda[3,1,1]",
  indirect_var = "TNIE_mda[3,1,1]",
  total_var    = "TE_mda[3,1,1]",
  effect_units = "log MDA"
)

plot_effect_decomposition(
  model_bb_base$sample,
  direct_var   = "PNDE_water[3,1,1]",
  indirect_var = "TNIE_water[3,1,1]",
  total_var    = "TE_water[3,1,1]",
  effect_units = "log water content"
)
```

